Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2022-11-20 latex/comp_or.tex
Parallel algorithms almost ready

We also developed Algorithm \ref{alg:greedy}, which generates a greedy allocation of the items available in node $l_k$, according to the non-ascending order of $\theta^k_{ij}$, and considering the consolidated items already shipped. Edges are included in this allocation as long as they respect feasibility constraints. Furthermore, the volume of each pallet $p_i$\/ cannot exceed $p_i.v \times limit$, where $ 0 < limit \leq 1$\/ is a parameter.

	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State $E^N_k \gets \varnothing$ 			
		\For{$i \gets 1$ to $m$}
			\State $pv[i] \gets 0$ 		
		\EndFor
		\For{$q \gets 1$ to $m_k$}
			\If{$(p_i, a_q^k) \in E^Q_k$} 
				\State $pv[i] \gets pv[i] + a_q^k.v$ 
			\EndIf		
		\EndFor		
		\For {each $e_{ij}^k$ in non-ascending order of $\theta^k_{ij}$}
			\If{($E^N_k \cup \{e_{ij}^k\}$ is feasible) {\bf and} ($pv[i] \leq p_i.v \times limit$)} 
				\State $E^N_k \gets E^N_k \cup \{e_{ij}^k\}$ 
				\State $pv[i] \gets pv[i] + t_j^k.v$ 	
			\EndIf
		\EndFor
		\State ${\bf return}\ G(V_k, E^N_k \cup E^Q_k)$

@@

	\caption{ $SolveNode(mpACO, k, G, numProcs)$ } \label{alg:aco}
	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State $G^*(V_k, E^{N*}_k \cup E^Q_k) \gets Greedy(k, G, 1)$  \label{aco:greedy}
		\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^*(V_k,E^{N*}_k \cup E^Q_k)$
		\State $A \gets numProcs$ \label{aco:numProcs}
		\State $ga^k_{ij}   \gets 0$, $1 \leq i \leq m$, $1 \leq j \leq n_k$  \label{aco:init1}
		\State $\tau^k_{ij} \gets 1$, $1 \leq i \leq m$, $1 \leq j \leq n_k$   \label{aco:init2}
		\State $\alpha    \gets 1.0$ \label{aco:alpha}
		\State $\beta     \gets 4.0$ \label{aco:beta}
		\State $\rho      \gets 0.2$ \label{aco:rho}
		\State $stagnant \gets 0$
		\While{ $stagnant < 3$ } \label{aco:while}

@@

A greedy solution is obtained in line \ref{aco:greedy}, being initially stored in $G^*$\/ and $G^+$.
The number $A$\/ of ants was set to the number of parallel processes (line \ref{aco:numProcs}).

@@

@@

In this heuristic, described in Algorithm \ref{alg:shims}, pallets $p_i$\/ are considered in non-descending order of $|p_i.d|$ (line \ref{shims:pallets}). For each $p_i$, the $n_k$\/ possible edges $e_{ ij}^k$\/ are considered in non-increasing order of $\theta_{ij}^k$\/ (line \ref{shims:edges}) in three phases: the greedy phase (lines \ref{shims:phase1a}-\ref{shims:phase1b}), the composition phase of the shims (lines \ref{shims:phase2a}-\ref{shims:phase2b}) and the selection phase of the best shim (lines \ref{shims:phase3a}-\ref{shims:phase3b}).

Figure \ref{fig:whip} represents the $n_k$\/ possible edges $e_{ij}^k$\/ of $p_i$\/ sorted by $\theta_{ij}^k$. \emph{mpShims}\/ starts with a greedy solution, stopping at the edge with index $\eta_1[i]$ close to the local optimum (first phase). Then, considering even the edge with the index $\eta_2[i]$, it elaborates different possible complements for this pallet (second phase) and selects the best of these complements (third phase).

\begin{algorithm}[H]
	\caption{ $SolveNode(mpShims, k, G, numProcs)$}  \label{alg:shims}
	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State Sort $M$ by $|p_i.d|$ in non-descending order \label{shims:pallets}
		\State $E^N_k \gets \varnothing$ 
		\State Let $E[1..m][1..n_k]$, where $E[i]$ is an array with $n_k$ edges $(p_i,t^k_j)$ sorted by $\theta_{ij}^k$ in non-ascending order, $1 \leq i \leq m$ \label{shims:edges}

@@

		\For{$i \gets 1$ to $m$} 
			\State $pv[i] \gets 0$		\label{shims:phase1a}
			\State $\eta_1[i] \gets 0$
			\While{($p_i.v - pv[i] > gap*p_i.v$) {\bf and} ($\eta_1[i]<n_k$)} \label{shims1:slack}
				\State $\eta_1[i] \gets \eta_1[i] + 1$ \label{shims:eta1}
				\State $e^k_{ij} \gets E[i][\eta_1[i]]$
				\If{($e^k_{ij}$ is feasible) {\bf and} ($t^k_j.v \leq p_i.v - pv[i]$)} 
					\State $E^N_k \gets E^N_k \cup \{e^k_{ij}\}$
					\State $pv[i] \gets pv[i] + t^k_j.v$ 
				\EndIf
			\EndWhile 
			\State $slack[i] \gets p_i.v - pv[i]$
			\State $\eta_2[i] \gets \eta_1[i]$
			\Repeat
				\State $\eta_2[i] \gets \eta_2[i] + 1$  \label{shims:eta2}
				\State $e^k_{ij} \gets E[i][\eta_2[i]]$
				\State $pv[i] \gets pv[i] + t^k_j.v$
			\Until{($\eta_2[i] = n_k$) {\bf or} ($ pv[i] \geq (1 + 3*gap)*p_i.v)$)} \label{shims:phase1b}

\State $Set \gets \varnothing$  \label{shims:phase2a}
\State $q \gets 1$
\State $shim[q] \gets \varnothing$
\State $volume \gets 0$
\For{$x \gets \eta_1[i]$ to $\eta_2[i]$}
\State $e_{ij}^k \gets E[i][x]$
\If {($e_{ij}^k \not\in E^N_k \cup shim[q]$) {\bf and} ($e_{ij}^k$ is feasible)}
\If{$t_j^k.v + volume \leq slack[i]$} \label{bincapacity}
\State $shim[q] \gets shim[q] \cup \{e_{ij}^k\}$	
\State $volume \gets volume + t_j^k.v$	
\Else \label{newbins}
\State $Set \gets Set \cup shim[q] $
\State $q \gets q+1$
\State $shim[q] \gets \varnothing$
\State $volume \gets 0$
\If{$t_j^k.v \leq slack[i]$} 
\State $shim[q] \gets shim[q] \cup \{e_{ij}^k\}$	
\State $volume \gets volume + t_j^k.v$
\EndIf
\EndIf
\EndIf
\EndFor   \label{shims:phase2b}

			\State $sh_w \gets S$, where $S \in Set$ and $\sum_{e_{ab}^k \in S} t_b^k.w$ is maximum \label{shims:phase3a}
			\State $sh_v \gets S$, where $S \in Set$ and $\sum_{e_{ab}^k \in S} t_b^k.v$ is maximum
			\State $sh_{best} \gets S$, where $S \in \{sh_w, sh_v\}$ and $\sum_{e_{ab}^k \in x} t_b^k.s$ is maximum
			\State $E^N_k \gets E^N_k \cup sh_{best}$ \label{shims:phase3b}
		\EndFor
		\State ${\bf return}$ $G(V_k, E^N_k \cup E^Q_k)$
	\end{algorithmic}
\end{algorithm}

In the first phase (lines \ref{shims:phase1a}-\ref{shims:phase1b}), a greedy and partial solution for $p_i$\/ is constructed by edges inclusion, until a certain slack or clearance is reached. This slack was empirically defined as $gap * p_i.v$\/ (line \ref{shims1:slack}), where $gap = 2000/n_k*m$\/ (line \ref{shims:gap}).

In the second phase (lines \ref{shims:phase2a}-\ref{shims:phase2b}), a set of shims named $Set$\/ is created, where each shim is formed by a group of edges in the range $[\eta_1[i],\eta_2[i]]$, whose total volume is limited by $slack[i]$. In this phase, the heuristic that provided the best results, both in terms of time and quality, is based on {\it Next-Fit}, which is an approximation algorithm for the {\it Bin Packing Problem}\/ \cite{JohnsonGarey1985}. Basically, shims are created by accumulating the following edges, taking $slack[i]$\/ as a limit.

In the third phase (lines \ref{shims:phase3a}-\ref{shims:phase3b}), the best shim in $Set$\/ is chosen. Initially, two shims are found: $sh_w$\/ with larger weight and $sh_v$\/ with larger volume. Among the two, the shim with the highest score will be chosen and its edges are inserted into $E^N_k$.

>>>>

@@

	\caption{ $SolveNode(mpACO, k, G, numProcs)$ } \label{alg:aco}
	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State $G^*(V_k, E^{N*}_k \cup E^Q_k), \_ \gets Greedy(k, G, 0.9)$  \label{aco:greedy}
		\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^*(V_k,E^{N*}_k \cup E^Q_k)$
		\State $A \gets numProcs$ \label{aco:numProcs}
		\State $ga^k_{ij}   \gets 0$, $1 \leq i \leq m$, $1 \leq j \leq n_k$  \label{aco:init1}
		\State $\tau^k_{ij} \gets 1$, $1 \leq i \leq m$, $1 \leq j \leq n_k$   \label{aco:init2}
		\State $\alpha    \gets 1.0$ \label{aco:alpha}
		\State $\beta     \gets 4.0$ \label{aco:beta}
		\State $\rho      \gets 0.2$ \label{aco:rho}
		\State $stagnant \gets 0$
		\While{ $stagnant < 3$ } \label{aco:while}

@@

A greedy solution is obtained in line \ref{aco:greedy}, being initially stored in $G^*$\/ and $G^+$. The argument 0.9 generates an initial and poor solution for ACO to improve. The second return value denoted by underscore is not used by ACO.

The number $A$\/ of ants was set to the number of parallel processes (line \ref{aco:numProcs}).

@@

We also developed Algorithm \ref{alg:greedy}, which generates a greedy allocation of the items available in node $l_k$, according to the non-ascending order of $\theta^k_{ij}$, and considering the consolidated items already shipped. Edges are included in this allocation as long as they respect feasibility constraints. Furthermore, the volume of each pallet $p_i$\/ cannot exceed $p_i.v \times limit$, where $ 0 < limit \leq 1$\/ is a given parameter.

	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$

		\State $\eta_1 \gets \{0\} \times m$ \label{greedy:eta}
		\State $volume \gets \{0\} \times m$

		\State $E^N_k \gets \varnothing$

		\For{$q \gets 1$ to $m$}
			\If{$(p_i, a_q^k) \in E^Q_k$} 
				\State $volume[i] \gets volume[i] + a_q^k.v$ 
			\EndIf		
		\EndFor

		\For {each $e_{ij}^k$ in non-ascending order of $\theta^k_{ij}$}
			\If{($E^N_k \cup \{e_{ij}^k\}$ is feasible) {\bf and} ($volume[i] \leq p_i.v \times limit$)} 
				\State $E^N_k \gets E^N_k \cup \{e_{ij}^k\}$ 
				\State $volume[i] \gets volume[i] + t_j^k.v$
				\State $\eta_1[i] \gets \eta_1[i] + 1$ 	
			\EndIf
		\EndFor
		\State ${\bf return}\ G(V_k, E^N_k \cup E^Q_k),\ \eta_1$

In Algorithm \ref{alg:greedy}, line \ref{greedy:eta} defines an array with the same size as pallets array to save the set of last edges inserted in the greedy solution.

@@

In this heuristic, described in Algorithm \ref{alg:shims}, pallets $p_i$\/ are considered in non-descending order of $|p_i.d|$ (line \ref{shims:pallets}).

Figure \ref{fig:whip} represents the $n_k$\/ possible edges $e_{ij}^k$\/ of $p_i$\/ sorted by $\theta_{ij}^k$. \emph{mpShims}\/ starts with a greedy solution, stopping at the edge with index $\eta_1$ close to the local optimum (first phase). Then, considering even the edge with the index $\eta_2[i]$, it elaborates different possible complements for this pallet (second phase) and selects the best of these complements (third phase).

\begin{algorithm}[H]
	\caption{ $SolveNode(mpShims, k, G, numProcs)$}  \label{alg:shims}
	\begin{algorithmic}[1]

		\State $G^* \gets G(V_k, E^Q_k)$

		\State Sort $M$ by $|p_i.d|$ in non-descending order \label{shims:pallets}

		\State $E^N_k \gets \{\}$

		\State Let $E[1..m][1..n_k]$, where $E[i]$ is an array with $n_k$ edges $(p_i,t^k_j)$ sorted by $\theta_{ij}^k$ in non-ascending order, $1 \leq i \leq m$ \label{shims:edges}

@@

		\State $shimsQueue \gets \{\}$, a queue of shims solutions to be used by the child processes

		\For{$lim \in limits$}
			\State $shimsSolve(G(V_k, E^Q_k),\ M,\ shimsQueue,\ lim,\ E)$
		\EndFor

		\State $counter \gets |limits|$

		\While{$runtime < 0.7s$ {\bf and} $counter > 0$}

			\State $G \gets shimsQueue.Pop()$	
			\If{$f_s(G) > f_s(G^*)$}
				\State $G^* \gets G$ 
			\EndIf

		\State ${\bf return}$ $G^*(V_k, E^N_k \cup E^Q_k)$
	\end{algorithmic}
\end{algorithm}

In this phase, the heuristic that provided the best results, both in terms of time and quality, is based on {\it Next-Fit}, which is an approximation algorithm for the {\it Bin Packing Problem}\/ \cite{JohnsonGarey1985}. Basically, shims are created by accumulating the following edges, taking $slack[i]$\/ as a limit.

\begin{algorithm}[H]
	\caption{ $shimsSolve(G,\ M,\ shimsQueue,\ limit,\ E)$}  \label{alg:shimsSolve}
	\begin{algorithmic}[1]

		\State $G(V_k,\ E^Q_k \cup E^N_k),\ \eta_1 \gets Greedy(k,\ G(V_k,\ E^Q_k),\ limit)$ \label{first_phase}

		\For{$i \gets 1$ to $m$}

			\State $\eta_2[i] \gets \eta_1[i]$

			\State $volume \gets 0$

			\Repeat \label{second_phase_begin}
				\State $\eta_2[i] \gets \eta_2[i] + 1$ 
				\State $e^k_{ij} \gets E[i][\eta_2[i]]$
				\State $volume \gets volume + t^k_j.v$
			\Until{($\eta_2[i] = n_k$) {\bf or} ($ volume \geq (2 - limit)*p_i.v)$)} \label{second_phase_end}

			\State $slack[i] \gets (1-limit) \times p_i.v$

			\State $E^N_k \gets E^N_k \cup getBestShims(i,\ \eta_1,\ \eta_2,\ E,\ k,\ slack[i])$ \label{third_phase}

		\State $shimsQueue \gets shimsQueue \cup \{G(V_k, E^N_k \cup E^Q_k)\}$ \label{enqueue}

Algorithm \ref{alg:shimsSolve}, line \ref{first_phase} corresponds to Shims' first phase. Lines  \ref{second_phase_begin}-\ref{second_phase_end} correspond to the Shims' second phase.
Line \ref{third_phase} corresponds to Shims' third phase. In the last line, \ref{enqueue}, {\it mpShims} puts the solution in the queue to be retrieved by the main process.

\begin{algorithm}[H]
	\caption{ $getBestShims(i,\ \eta_1,\ \eta_2,\ E,\ k,\ slack)$ }  \label{alg:getBestShims}
	\begin{algorithmic}[1]

		\State $volume \gets 0$

		\State $b \gets 1$
		\State $shims[b] \gets \{\}$
		\State $Set \gets Set \cup \{shims[b]\}$

		\For{$x \gets \eta_1$ to $\eta_2[i]$}

			\State $NewShims \gets$ {\bf True}
			\State $e_{ij}^k \gets E[i][x]$

			\For{$shims \in Set$}

				\If { $e_{ij}^k \not\in (E^N_k \cup shims)$ {\bf and} $e_{ij}^k$ is feasible  {\bf and}  $(t_j^k.v + volume) \leq slack[i]$}

					\State $shims \gets shims \cup \{e_{ij}^k\}$
					\State $volume \gets volume + t_j^k.v$
					\State $NewShims \gets$ {\bf False}

					\State {\bf break}

				\EndIf

			\If{$NewShims$} \label{ffd:new1}
				\State $volume \gets 0$
				\State $b \gets b + 1$
				\State $shims[b] \gets \{\}$
				\State $shims[b] \gets shims[b] \cup \{e_{ij}^k\}$
				\State $Set \gets Set \cup \{shims[b]\}$
			\EndIf

		\State $sh_w \gets shims$, where $shims \in Set$ and $\sum_{e_{ab}^k \in shims} t_b^k.w$ is maximum 		
		\State $sh_v \gets shims$, where $shims \in Set$ and $\sum_{e_{ab}^k \in shims} t_b^k.v$ is maximum
		\State $sh_{best} \gets shims$, where $shims \in \{sh_w, sh_v\}$ and $\sum_{e_{ab}^k \in x} t_b^k.s$ is maximum

		\State ${\bf return}$ $sh_{best}$
